Criticisms of the IHME COVID-19 model

### 2020-04-27 text/20/ihme-covid-model.Rmd
Created data-repositories.md

@@

You will never resolve the controversy over theoretical versus atheoretical models, but in infectious disease modeling, the theoretical models have several important advantages.

First, the SEIR models are battle-hardened. They have been used successfully many times in the past. Every year a new flu season offers a new test of these models and every year they seem to do reasonably well.

@@

@@

@@

>>>>

There are simulation models that are very close in spirit to the SEIR models, but which require a much greater degree of computational power. These simulations create random interactions among thousands or millions of individuals where each interaction has a specified probability of spreading the disease during an interaction between an infected individual and a susceptible individual. Other than the far greater amount of computing power needed, these models have the basic features as the SEIR models.

@@

You will never resolve the controversy over theoretical versus atheoretical models, but in infectious disease modeling, the theoretical models have several important advantages. 
First, the SEIR models are battle-hardened. They have been used successfully many times in the past. Every year a new flu season offers a new test of these models and every year they seem to do reasonably well.

@@

Fourth, the theoretical models allow you to simulate scenarios that have not yet occured, such as what happens when we introduce a vaccine or whether a lack of permanent immunity after recovery will lead to a new wave of infections.

@@

Chow D. What we know about the coronavirus model the White House unveiled. NBC News, 2020-04-01. Available in [html format](https://www.nbcnews.com/science/science-news/what-we-know-about-coronavirus-model-white-house-unveiled-n1173601).

+ This article was also written in response to that White House press briefing. It emphasizes the many sources of uncertainty in these predictions.

@@

Chowell G, Sattenspiel L, Bansal S, Viboud C. Mathematical models to characterize early epidemic growth: A review. Phys Life Rev. 2016;18:66–97. doi:10.1016/j.plrev.2016.07.005. Available in [html format](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC5348083/) or [pdf format](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC5348083/pdf/nihms-802148.pdf).

+ This article was written before the COVID=19 pandemic. It provides a nice review of how SEIR models work and applies them to a variety of pandemics.

Kissler SM, Tediganto C, Goldstein E, Grad YH, Lipsitch M. Projecting the transmission dynamics of SARS-CoV-2 through the postpandemic period. Science 2020-04-14. DOI: 10.1126/science.abb5793, Available in [html format](https://science.sciencemag.org/content/early/2020/04/24/science.abb5793).

+ This article compares SARS-CoV-2 (the virus that causes COVID-19) to previous viruses to assess the chances that we will see future seasonal waves of COVID-19. This is an example of the type of modeling that can only be done with a theoretical model.

FRED Modeling System. Available in [html format](https://fred.publichealth.pitt.edu/).

+ FRED stands for Framework for Reconstructing Epidemic Dynamics and represents a software system that uses simulation to model epidemics. It has been used to model outbreaks of measles and influenza, but not yet (as far as I can tell) for COVID-19. It is available for anyone to download and use on their own.

### Acknowledgments

Thanks to Chris Barker and Ruth Etzioni for helpful suggestions.
